---
title: "Final_Workspace"
format: html
editor: visual
---

# Final project polygons work

```{r}
library(sf)
library(leaflet)
library(htmltools)
library(glue)
```

```{r}
final_parks <- read_csv("final_parks.csv") |>
  mutate(name = case_when(
    name == "Cascade River State Park Recreation Site" ~ "Cascade River State Park",
    name == "Killen Woods State Park" ~ "Kilen Woods State Park",
    name == "John Latsch State Park" ~ "John A Latsch State Park",
    name == "Nerstrand Woods State Park" ~ "Nerstrand Big Woods State Park",
    name == "Sakatah State Park" ~ "Sakatah Lake State Park",
    name == "O L Kipp State Park" ~ "Great River Bluffs State Park",
    name == "Saint Croix Wild River State Park" ~ "Wild River State Park",
    name == "Helmer Myre State Park" ~ "Myre-Big Island State Park",
    .default = name
    ))

park_ids <- read_csv("Dnr_Stat_Plan_Areas_Prk.csv")

park_polygons <- read_sf("Dnr_Stat_Plan_Areas_Prk.geojson") |>
  mutate(legislative_name = str_replace_all(legislative_name, "\\.", ""),
         legislative_name = ifelse(legislative_name == "St Croix State Park", "Saint Croix State Park", legislative_name))
```

extras: - kadunce river - upper sioux

# missing parks -- finding data

-   crow wing state park
-   forestville
-   franz jevne
-   glendalough
-   grand portage
-   lac qui parle
-   lake louise
-   lake shetek
-   lake vermilion
-   monson lake
-   moose lake
-   rice lake
-   savanna portage
-   schoolcraft

```{r}
name <- c("Crow Wing", "Forestville Mystery Cave", "Franz Jevne", "Glendalough", "Grand Portage", "Lac qui Parle", "Lake Louise", "Lake Shetek", "Lake Vermilion-Soudan Underground Mine", "Monson Lake", "Moose Lake", "Rice Lake", "Savanna Portage", "Schoolcraft", "Cascade River", "Father Hennepin", "Myre-Big Island", "Great River Bluffs", "Wild River")

web_id <- c(139, 148, 157, 167, 173, 197, 214, 220, 285, 238, 239, 250, 259, 263, 133, 142, 175, 244, 254)

# checking scraping allowed
robotstxt::paths_allowed("https://www.dnr.state.mn.us/")

# checking conditions
session <- bow("https://www.dnr.state.mn.us/state_parks/list_alpha.html", force = TRUE)

# making list
urls <- vector("character")
for (i in 1:length(web_id)) {
  url <- str_c("https://www.dnr.state.mn.us/state_parks/park.html?id=spk00", web_id[i], "#information")
  urls <- c(urls, url)
}

# scraping each page
scrape_park_info <- function(url) {
  Sys.sleep(5)
  session <- bow(url, force = TRUE)
  scrape(session) |>
    html_nodes("meta[name='dc.description']") |>
    html_attr("content")
}

# building outside of the function
Sys.sleep(5)
  session <- bow(urls[1], force = TRUE)
  scrape(session) |>
    html_nodes("meta[name='dc.description']") |>
    html_attr("content")

# getting descriptions
description <- vector("character")
for (i in 1:length(urls)) {
  park_url <- urls[i]
  paragraph <- scrape_park_info(park_url)
  description <- c(description, paragraph) 
}

missing <- tibble(name, description) |>
  mutate(name = str_c(name, " State Park"))

new <- final_parks |>
  #left_join(missing, by = "name") |>
  full_join(missing, by = "name") |>
  mutate(description = coalesce(description.x, description.y),
         type = ifelse(is.na(type), "state park", type)) |>
  select(name, url, address, description, type)

#making a csv
write_csv(new, "parks_updated.csv")
```

# loading in updated parks csv

```{r}
parks <- read_csv("parks_updated.csv")
```

# added html labels, fixed for extra parks

```{r}
parks_sf <- parks |>
  left_join(park_polygons, by = c("name" = "legislative_name"))

parks_sf <- parks_sf |>
  mutate(labels = str_c("<p><strong>", name, "</strong></p>
<p><span style='font-weight: 400;'>", description, "</span></p>"))

labels <- lapply(parks_sf$labels, HTML)

parks_sf <- st_as_sf(parks_sf)

class(parks_sf)

leaflet(parks_sf) |>
  setView(-93.8, 46.5, 6) |>
  addTiles() |>
  addPolygons(
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 1,
    weight = 2,
    opacity = 1,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "red",
      bringToFront = TRUE),
    label = labels,
    labelOptions = popupOptions( 
      style = list("font-weight" = "normal", padding = "2px 5px", color = "black", "width" = "450px", "white-space" = "normal"),
      textsize = "15px",
      direction = "auto")
  )
```

# scraping wikipedia

```{r}
session <- bow("https://en.wikipedia.org/wiki/List_of_Minnesota_state_parks", force = TRUE)

result <- scrape(session) |>
  html_nodes(css = "table") |>
  html_table(header = TRUE, fill = TRUE)
try <- result[[1]]
try <- try[-1,] 
  
```
