---
title: "Kaisa"
format: html
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: FALSE

library(tidyverse)
library(rvest)
library(polite)
library(dplyr)
```


## Getting lists of parks
```{r}
robotstxt::paths_allowed("https://stateparks.com/")


session <- bow("https://stateparks.com/minnesota_parks_and_recreation_destinations.html", force = TRUE)

#better version?
Sys.sleep(5)
parks_text <- scrape(session) |>
  html_nodes("#parklink a") |>
  html_text()

#mn_parks <- read_html("https://stateparks.com/minnesota_parks_and_recreation_destinations.html")

#parks_list <- html_nodes(mn_parks, "#parklink a")

#parks_text <- html_text(parks_list)
```

## test - writing a formula
```{r}
form <- parks_text[116] |>
  str_to_lower() |>
  str_replace_all(c(" " = "_", "'" = "_", "wildlife_mgmt." = "state_wildlife_management"))

url <- str_c("https://stateparks.com/", form, "_in_minnesota.html")

print(url_list)

# formula -- generates urls we can scrape (only in mn) --------------------------
url_list <- vector("character")
for (i in 1:length(parks_text)) {
  form <- parks_text[i] |>
  str_to_lower() |>
  str_replace_all(c(" " = "_", "'" = "", "wildlife_mgmt." = "state_wildlife_management"))
  url <- str_c("https://stateparks.com/", form, "_in_minnesota.html")
  url_list <- c(url_list, url)
}

# I just updated this so that all the urls are fixed! some weren't correct before
```




## scraping the first paragraph

```{r}
# function using bow/scrape for finding paragraphs in parks 
scrape_page <- function(url) {
  Sys.sleep(5) # this follows the website robots requests
  session <- bow(url, force = TRUE) #setting session
  scrape(session) |>
    html_nodes("#overview :nth-child(5)") |>
    html_text()
}

# function using bow/scrape for finding paragraphs in parks (option 2)
scrape_page2 <- function(url) {
  Sys.sleep(5)
  session <- bow(url, force = TRUE)
  scrape(session) |>
    html_nodes("#overview .psbod") |>
    html_text()
}

# i think both of these work? I used the first but we could probably pick either


# scraping the parks -----------------------------------------------------------
paragraph_list <- vector("character")
for (i in 1:length(url_list)) {
  park_url <- url_list[i]
  paragraph <- scrape_page(park_url)
  paragraph_list <- c(paragraph_list, paragraph) 
}

print(paragraph_list) 

#cleaning the paragraph list-- not perfect, it removes both extra empty strings AND those indicating no paragraph was available

park_par_tbl <- paragraph_list |>
  str_subset("[a-zA-Z]+") |>
  str_subset("Email:|Picnickingyes", negate = TRUE) |> 
  as_tibble() 

#MISSING cascade river, father hennepin, helmer myre, kadunce river, o l kipp, st croix wild river, boekholt grove, cooks, coon lake, little elk, malardi lake, talcot lake, teiken-dalve, 
```


## scraping addresses

```{r}
# function for scraping addresses
scrape_page3 <- function(url) {
  Sys.sleep(5)
  session <- bow(url, force = TRUE)
  scrape(session) |>
    html_nodes(".parkinfo div:nth-child(1)") |> # css selector for the address box,
    html_text()                                 # but it selects park name as well
}

# checking it works on individual urls
Sys.sleep(5)
session <- bow("https://stateparks.com/badoura_state_forest_in_minnesota.html", force = TRUE)
scrape(session) |>
  html_nodes(".parkinfo div:nth-child(1)") |>
    html_text()

# scraping addresses
address_list <- vector("character")
for (i in 1:length(url_list)) {
  park_url2 <- url_list[i]
  address <- scrape_page3(park_url2)
  address_list <- c(address_list, address) 
}

viewing <- address_list |>
  as_tibble()

# address list has a lot of extra characters that aren't visible in the tibble, and I haven't cleaned it yet
```

```{r}
park_names <- as_tibble(parks_text)

indexes <- c(9, 11, 20, 26, 37, 40, 107, 109, 110, 112, 113, 114, 115)

for (i in 1:13) {
  park_par_tbl <- append(park_par_tbl, "", after = indexes[i])
}

park_pars <- as_tibble(park_par_tbl)

new <- bind_cols(park_names, park_pars)
```

```{r}
cleaned_addresses <- address_list |>
  # replace line breaks and tabs with spaces
  str_replace_all("[\r\n\t]", " ") |>
  #replace multiple spaces with a single space
  str_replace_all("  +", " ") |>
  #remove email addresses
  str_replace("Email:.*$", "") |>
  #remove phone numbers  
  str_replace("Phone:.*$", "") |>
  #remove reservation info
  str_replace("Reservations:.*$", "") |>
  str_replace_all("^[^\\d]*", "") # removes the extra park name at the front



clean2 <- cleaned_addresses |>
  str_replace_all("^[^\\d]*", "")

# same as Romina's code, but i fixed the paragraphs (my bad originally!)
final <- tibble(
  name = parks_text,
  url = url_list,
  address = cleaned_addresses,
  description = park_par_tbl,
  nature = nature_list2
) 

# replaced all the blanks with NAs
final_parks <- final %>% mutate_all(na_if,"")

#the final tidy tibble
#final_parks <- tibble(
 # park_name = parks_text,
  #park_url = url_list,
  #address = cleaned_addresses,
  #description = park_par_tbl$description[1:length(parks_text)]
#)

#the first few rows
head(final_parks)

#the structure
str(final_parks)

# missing data
#final_parks |>
 # summarise(
  #  total_parks = n(),
  #  parks_with_address = sum(!is.na(address) & address != ""),
  #  parks_with_description = sum(!is.na(description) & description != "")
 # )

```



# optional scraping nature (and some history) sections
```{r}
scrape_page_nature <- function(url) {
  Sys.sleep(5)
  session <- bow(url, force = TRUE)
  scrape(session) |>
    html_nodes(".psbod:nth-child(8)") |>
    html_text()
}

nature_list <- vector("character")
for (i in 1:length(url_list)) {
  park_url2 <- url_list[i]
  nature <- scrape_page_nature(park_url2)
  nature_list <- c(nature_list, nature) 
}

view(as_tibble(nature_list)) # missing: cascade river, father hennepin, helmer myre, 19(interstate) is the history, kadunce river, 27/28(lake carlos/maria) is history, ol kipp, st croix wild river, all forest except crow wing and grad portage, all mgmt except lac qui parle, all others except minnesota valley

park_names <- as_tibble(parks_text)

indexes <- c(9, 11, 20, 26, 37, 40, 52:63, 65:72, 74:110, 112:120, 122:124)

nature_list2 <- nature_list

for (i in 1:75) {
  nature_list2 <- append(nature_list2, "", after = indexes[i])
}

view(as_tibble(nature_list2))
```

# adding park type
```{r}
final_parks <- final_parks |>
  mutate(type = case_when(
    str_detect(parks_text, "State Park") ~ "state park",
    str_detect(parks_text, "State Forest") ~ "state forest",
    str_detect(parks_text, "Mgmt. Area") ~ "wildlife management area",
    str_detect(parks_text, "National Park") ~ "national park",
    str_detect(parks_text, "National Forest") ~ "national forest",
    str_detect(parks_text, "Wildlife Refuge") ~ "wildlife refuge"
  ))




str_detect(parks_text, "State Park")
str_detect(parks_text, "State Forest")
str_detect(parks_text, "Mgmt. Area")
str_detect(parks_text, "National Park")
str_detect(parks_text, "National Forest")
str_detect(parks_text, "Wildlife Refuge")
```

# Final project polygons work
```{r}
library(sf)
library(leaflet)
library(htmltools)
library(glue)
```

```{r}
final_parks <- read_csv("final_parks.csv") |>
  mutate(name = case_when(
    name == "Cascade River State Park Recreation Site" ~ "Cascade River State Park",
    name == "Killen Woods State Park" ~ "Kilen Woods State Park",
    name == "John Latsch State Park" ~ "John A Latsch State Park",
    name == "Nerstrand Woods State Park" ~ "Nerstrand Big Woods State Park",
    name == "Sakatah State Park" ~ "Sakatah Lake State Park",
    name == "O L Kipp State Park" ~ "Great River Bluffs State Park",
    name == "Saint Croix Wild River State Park" ~ "Wild River State Park",
    name == "Helmer Myre State Park" ~ "Myre-Big Island State Park",
    .default = name
    ))

park_ids <- read_csv("Dnr_Stat_Plan_Areas_Prk.csv")

park_polygons <- read_sf("Dnr_Stat_Plan_Areas_Prk.geojson") |>
  mutate(legislative_name = str_replace_all(legislative_name, "\\.", ""),
         legislative_name = ifelse(legislative_name == "St Croix State Park", "Saint Croix State Park", legislative_name))
```

extras: 
- kadunce river
- upper sioux


# missing parks -- finding data
- crow wing state park
- forestville
- franz jevne
- glendalough
- grand portage
- lac qui parle
- lake louise
- lake shetek
- lake vermilion
- monson lake
- moose lake
- rice lake
- savanna portage
- schoolcraft

```{r}
name <- c("Crow Wing", "Forestville Mystery Cave", "Franz Jevne", "Glendalough", "Grand Portage", "Lac qui Parle", "Lake Louise", "Lake Shetek", "Lake Vermilion-Soudan Underground Mine", "Monson Lake", "Moose Lake", "Rice Lake", "Savanna Portage", "Schoolcraft", "Cascade River", "Father Hennepin", "Myre-Big Island", "Great River Bluffs", "Wild River")

web_id <- c(139, 148, 157, 167, 173, 197, 214, 220, 285, 238, 239, 250, 259, 263, 133, 142, 175, 244, 254)

# checking scraping allowed
robotstxt::paths_allowed("https://www.dnr.state.mn.us/")

# checking conditions
session <- bow("https://www.dnr.state.mn.us/state_parks/list_alpha.html", force = TRUE)

# making list
urls <- vector("character")
for (i in 1:length(web_id)) {
  url <- str_c("https://www.dnr.state.mn.us/state_parks/park.html?id=spk00", web_id[i], "#information")
  urls <- c(urls, url)
}

# scraping each page
scrape_park_info <- function(url) {
  Sys.sleep(5)
  session <- bow(url, force = TRUE)
  scrape(session) |>
    html_nodes("meta[name='dc.description']") |>
    html_attr("content")
}

# building outside of the function
Sys.sleep(5)
  session <- bow(urls[1], force = TRUE)
  scrape(session) |>
    html_nodes("meta[name='dc.description']") |>
    html_attr("content")

# getting descriptions
description <- vector("character")
for (i in 1:length(urls)) {
  park_url <- urls[i]
  paragraph <- scrape_park_info(park_url)
  description <- c(description, paragraph) 
}

missing <- tibble(name, description) |>
  mutate(name = str_c(name, " State Park"))

new <- final_parks |>
  #left_join(missing, by = "name") |>
  full_join(missing, by = "name") |>
  mutate(description = coalesce(description.x, description.y),
         type = ifelse(is.na(type), "state park", type)) |>
  select(name, url, address, description, type)

#making a csv
write_csv(new, "parks_updated.csv")
```

# loading in updated parks csv
```{r}
parks <- read_csv("parks_updated.csv")
```

# added html labels, fixed for extra parks
```{r}
parks_sf <- parks |>
  left_join(park_polygons, by = c("name" = "legislative_name"))

parks_sf <- parks_sf |>
  mutate(labels = str_c("<p><strong>", name, "</strong></p>
<p><span style='font-weight: 400;'>", description, "</span></p>"))

labels <- lapply(parks_sf$labels, HTML)

parks_sf <- st_as_sf(parks_sf)

class(parks_sf)

leaflet(parks_sf) |>
  setView(-93.8, 46.5, 6) |>
  addTiles() |>
  addPolygons(
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 1,
    weight = 2,
    opacity = 1,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "red",
      bringToFront = TRUE),
    label = labels,
    labelOptions = popupOptions( 
      style = list("font-weight" = "normal", padding = "2px 5px", color = "black", "width" = "450px", "white-space" = "normal"),
      textsize = "15px",
      direction = "auto")
  )
```


# scraping wikipedia
```{r}
session <- bow("https://en.wikipedia.org/wiki/List_of_Minnesota_state_parks", force = TRUE)

result <- scrape(session) |>
  html_nodes(css = "table") |>
  html_table(header = TRUE, fill = TRUE)
try <- result[[1]]
try <- try[-1,] 
try <- try[,-4]
try <- try |> rename(Acres = `Size[9]`)

# removes . and / and renames parks to help with joining with the parks csv
try <- try |>
  mutate(`Park name` = str_replace_all(`Park name`, "[.]", "")) |>
  mutate(`Park name` = str_replace_all(`Park name`, "[/]", " ")) |>
  mutate(`Park name` = case_when(
    `Park name` == "Nerstrand-Big Woods State Park" ~ "Nerstrand Big Woods State Park",
    `Park name` == "St Croix State Park" ~ "Saint Croix State Park",
    `Park name` == "Lake Vermilion Soudan Underground Mine State Park" ~ "Lake Vermilion-Soudan Underground Mine State Park",
    .default = `Park name`),
    Coordinates = str_replace_all(Coordinates, "^.*nowrap\\}", "")) |>
  mutate(Coordinates = str_replace_all(Coordinates, "^.*/ ", ""),
         Coordinates = str_replace_all(Coordinates, "\\(.*\\)", "")) |>
  separate_wider_delim(Coordinates, delim = ";", names = c("lat", "long")) |>
  mutate(long = str_trim(long), lat = str_trim(lat))
```

```{r}
#trying to convert to numeric
library(readr)
numeric <- try |>
  mutate(long = str_replace_all(long, "^[^\\d]*", "-")) |>
  mutate(lat = as.numeric(lat), long = parse_number(long))

numeric <- read_csv("wiki_parks_data.csv")
```


```{r}
county_parks <- numeric |>
  mutate(`County or Counties` = case_when(
    str_detect(`County or Counties`, "[a-z] and ") ~ str_replace_all(`County or Counties`, " and", ","),
    str_detect(`County or Counties`, "[a-z], and ") ~ str_replace_all(`County or Counties`, " and", ""),
    .default = `County or Counties`)) |>
  mutate(unpacked = str_split(`County or Counties`, ",")) |>
  filter(!is.na(unpacked)) |>
  unnest(cols = c(unpacked)) |>
  mutate(county = str_trim(unpacked)) |>
  #pivot_wider(names_from = county, values_from = n, values_fill = list(n = 0))
  group_by(county) |>
  mutate(num = n())
  
county_parks <- county_parks |>
  mutate(county = ifelse(county == "Murray County", "Murray", county))

```

# mapping
```{r}
county_polygons <- read_sf("Mn_County_Boundaries_Multipart.geojson")

county_maps <- county_parks |>
  right_join(county_polygons, by = c("county" = "countyname")) |>
  mutate(num = replace_na(num, 0))

pal1 <- colorFactor(c("white", "yellow2", "yellowgreen", "forestgreen", "cyan4", "blue4"), county_maps$num)

county_maps <- county_maps |>
  select(-geometry) |>
  mutate(labels = str_c(county, " County: ", num, ifelse(num == 1, " state park", " state parks")),
         labels2 = `name`)

labels <- lapply(county_maps$labels, HTML)
labels2 <- lapply(county_maps$labels2, HTML)

county_maps <- st_as_sf(county_maps)
class(county_maps)

leaflet(county_maps) |>
  setView(-93.8, 46.5, 6) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(num),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "white",
      dashArray = "",
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels2,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal, values = ~num, opacity = 0.7, title = NULL,
    position = "bottomright")
```

```{r}
southern <- list("Traverse", "Big Stone", "Swift", "Lac qui Parle", "Chippewa", "Kandiyohi", "Meeker", "Renville", "McLeod", "Sibley", "Nicollet", "Le Sueur", "Rice", "Goodhue", "Wabasha", "Winona", "Houston", "Fillmore", "Mower", "Olmsted", "Dodge", "Steele", "Freeborn", "Waseca", "Faribault", "Blue Earth", "Martin", "Watonwan", "Brown", "Cottonwood", "Jackson", "Nobles", "Murray", "Rock", "Lyon", "Redwood", "Yellow Medicine", "Lincoln", "Pipestone")

metro <- list("Dakota", "Scott", "Carver", "Wright", "Ramsey", "Hennepin", "Washington", "Chisago", "Anoka")

arrowhead <- list("Isanti", "Kanabec", "Aitkin", "Itasca", "Koochiching", "Pine", "Carlton", "St. Louis", "Lake", "Cook")

heartland <- list("Stevens", "Grant", "Douglas", "Pope", "Stearns", "Sherburne", "Benton", "Mille Lacs", "Crow Wing", "Cass", "Beltrami", "Lake of the Woods", "Roseau", "Kittson", "Marshall", "Pennington", "Clearwater", "Red Lake", "Polk", "Norman", "Mahnomen", "Clay", "Becker", "Hubbard", "Wadena", "Otter Tail", "Wilkin", "Todd", "Morrison")

county_maps <- county_maps |>
  mutate(region = case_when(county %in% southern ~ "Southern",
                            county %in% heartland ~ "Heartland",
                            county %in% metro ~ "Metro",
                            county %in% arrowhead ~ "Arrowhead",
                            .default = name)) |>
  group_by(region) |>
  mutate(count = n())


pal2 <- colorFactor(c("yellow2", "forestgreen","cyan4", "blue4"), county_maps$count)

county_maps <- county_maps |>
  mutate(labels = str_c(county, " County: ", num, ifelse(num == 1, " state park", " state parks")),
         labels2 = name,
         labels3 = str_c(region, " Region: ", count, "state parks"))

labels <- lapply(county_maps$labels, HTML)
labels2 <- lapply(county_maps$labels2, HTML)
labels3 <- lapply(county_maps$labels3, HTML)

county_maps <- st_as_sf(county_maps)
class(county_maps)

leaflet(county_maps) |>
  setView(-93.8, 46.5, 6) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(count),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal, values = ~count, opacity = 0.7, title = NULL,
    position = "bottomright")

county_maps |> write_sf("mapping_data.geojson")
```  
  


