---
title: "Minnesota State Parks"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "Github", href: "https://github.com/philli17/MiniProject2" }
    theme:
      version: 4
      bootswatch: sandstone
runtime: shiny
---

```{r}
library(leaflet)
library(tidyverse)
library(sf)
library(readr)
library(htmltools)
```

```{r, label = "reading in csvs", include = FALSE}
numbers <-read_csv("numbers.csv")

with_wiki <- read_csv("wiki_parks_data.csv")

parks <- read_csv("parks_updated.csv")

park_polygons <- read_sf("Dnr_Stat_Plan_Areas_Prk.geojson") |>
  mutate(legislative_name = str_replace_all(legislative_name, "\\.", ""),
         legislative_name = ifelse(legislative_name == "St Croix State Park", "Saint Croix State Park", legislative_name))

county_maps <- read_csv("mapping_data.csv")

county_polygons <- read_sf("Mn_County_Boundaries_Multipart.geojson")
```

About
==========================

Column {data-width = 200}
-------------------------

**About This Project** 

Minnesota is known as the Land of 10,000 Lakes, but its parks are equally impressive. For this project, we analyzed data in Minnesota’s state parks to understand what makes them unique and how visitation patterns vary across seasons and locations.

The result is an interactive Shiny dashboard featuring maps, plots, and data tables that let users explore Minnesota’s state parks in depth. Whether you’re planning a trip or researching park trends, this tool provides insight into the state’s natural spaces. 

**Data Sources:** [stateparks.com](http://stateparks.com), [Wikipedia](https://en.wikipedia.org/wiki/List_of_Minnesota_state_parks), [MN DNR GIS](https://gis.data.mn.gov/maps/a42128766db2447cb77a247d4a074173/about), [Joint Legacy Research Project](https://storymaps.arcgis.com/stories/8489886967dc48b58cb8f54e1a3c014f)

**Authors:** Mary Phillips, Kaisa Nilsson, Romina Amirian


Graphs
==========================

Column 1
--------------------------------------------------
### Visitation by Year Established 

This scatterplot compares visitation numbers to the year the park was established. It isn't the most useful stat, and it doesn't tell us much, but we can see that the oldest park (established in 1891, Itasca State Park) also has the highest number of visits, by far. As far as newer parks getting consistently lower visitation numbers, we probably can't draw that conclusion.

### Scatterplot
```{r}
renderPlot({
  ggplot(numbers, aes(x = Year,
                      y = total,
                      color = Season)) +
    geom_point() +
    labs(x = "Year Park Established", 
         y = "Total Visitation by Car",
         color = "Region",
         title = "Visitation by Year Established")
})
```

Column 2
--------------------------------------------------
### Visitation by Park Size (and Region)

We can also look at some other data. This scatterplot charts the number of visitors against the number of acres of a park. The dots are colored based on their location. There are four regions of state parks in Minnesota: Arrowhead (northeast), Heartland (northwest), Metro (Twin-Cities), and Southern. This isn't the strongest relationship for a scatterplot to show, but we can see that the biggest park is much bigger than the other parks (the biggest is over 30000 acres, and the next biggest is only about 15000, half the size!). We can also see that the biggets parks with the most visitation are in the north - the Arrowhead and Heartland regions, while most Southern and Metro area parks are much smaller and have gotten fewer visitors.

### Scatterplot
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(x = Acres,
                      y = total,
                      color = region)) +
    geom_point() +
    labs(x = "Size of Park (acres)", 
         y = "Total Visitation by Car",
         title = "Visitation by Park Size")
})
```

Graphs - Interactive!
==========================

Bar Chart Inputs {.sidebar}
-------------------------

```{r, echo = FALSE}
inputPanel(
  selectInput("boxx", label = "x-axis variable:",
              choices = c("Year of Visit" = "visit_year",
                          "Visitation" = "total",
                          "Season" = "Season",
                          "Park Size (acres)" = "Acres",
                          "County of Park" = "County or Counties",
                          "Year Established" = "Year",
                          "Region" = "region")),
  selectInput("boxy", label = "y-axis variable:",
              choices = c("Visitation" = "total",
                          "Year of Visit" = "visit_year",
                          "Season" = "Season",
                          "Park Size (acres)" = "Acres",
                          "County of Park" = "County or Counties",
                          "Year Established" = "Year",
                          "Region" = "region")),
  selectInput("color", label = "grouping variable:",
              choices = c("Season" = "Season",
                          "Year of Visit" = "visit_year",
                          "County of Park" = "County or Counties",
                          "Year Established" = "Year",
                          "Region" = "region"))
)
```

Feel free to change around the variables with this bar chart as well! Our suggestion is x = Year of Visit, y = Visitation, grouping variable = Season - it makes it easy to see that there's more visits during summer, but also more during fall than spring, and winter is predictably the least visitation. Explore other combinations as well!

Column
--------------------------
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(x = .data[[input$boxx]], y = .data[[input$boxy]], color = .data[[input$color]], fill = .data[[input$color]])) + 
    geom_col()
})
```

Maps - Individual Parks
==========================


```{r, label = "adding html labels", include = FALSE}
# joining polygons
parks_sf <- parks |>
  left_join(park_polygons, by = c("name" = "legislative_name"))

# adding html labels of name/description
parks_sf <- parks_sf |>
  mutate(labels1 = str_c("<p><strong>", name, "</strong></p>
<p><span style='font-weight: 400;'>", description, "</span></p>"))

# applying html
labels1 <- lapply(parks_sf$labels1, HTML)

# making the dataframe sf
parks_sf <- st_as_sf(parks_sf)

# checking class
class(parks_sf)
```

```{r}
renderLeaflet({
  leaflet(parks_sf) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 1,
    weight = 2,
    opacity = 1,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "red",
      bringToFront = TRUE),
    label = labels1,
    labelOptions = popupOptions( 
      style = list("font-weight" = "normal", padding = "2px 5px", color = "black", "width" = "450px", "white-space" = "normal"),
      textsize = "15px",
      direction = "auto")
  )
})
```

Maps - Counties
===================================

{.sidebar}
----------------------------------
Discussing the Map

```{r, label = "lat long stuff", include = FALSE}
county_maps <- county_maps |>
  right_join(county_polygons, by = c("county" = "countyname")) |>
  mutate(num = replace_na(num, 0))
```

```{r, label = "setting up map", include = FALSE}
pal1 <- colorFactor(c("white", "yellow2", "yellowgreen", "forestgreen", "cyan4", "blue4"), county_maps$num)

labels <- lapply(county_maps$labels, HTML)
labels2 <- lapply(county_maps$labels2, HTML)
labels3 <- lapply(county_maps$labels3, HTML)

county_maps <- st_as_sf(county_maps)
class(county_maps)
```

Column {data-height = 3000}
------------------------------------
```{r}
renderLeaflet({
  leaflet(county_maps) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal1(num),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "white",
      dashArray = "",
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels2,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal1, values = ~num, opacity = 0.7, title = NULL,
    position = "bottomright")
})
```

Maps - Regions
==============================

```{r}
pal2 <- colorFactor(c("yellow2", "forestgreen","cyan4", "blue4"), county_maps$count)
```

```{r}
renderLeaflet({
  leaflet(county_maps) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal2(count),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal2, values = ~count, opacity = 0.7, title = NULL,
    position = "bottomright")
})
```

