---
title: "Minnesota State Parks"
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bootswatch: sandstone
runtime: shiny
---

```{r}
library(leaflet)
library(tidyverse)
library(sf)
library(readr)
```

```{r, label = "reading in csvs", include = FALSE}
numbers <-read_csv("numbers.csv")

with_wiki <- read_csv("wiki_parks_data.csv")

parks <- read_csv("parks_updated.csv")

park_polygons <- read_sf("Dnr_Stat_Plan_Areas_Prk.geojson") |>
  mutate(legislative_name = str_replace_all(legislative_name, "\\.", ""),
         legislative_name = ifelse(legislative_name == "St Croix State Park", "Saint Croix State Park", legislative_name))
```

About
==========================

Column {data-width = 200}
-------------------------

**About This Project** 

Minnesota is known as the Land of 10,000 Lakes, but its parks are equally impressive. For this project, we analyzed data in Minnesota’s state parks to understand what makes them unique and how visitation patterns vary across seasons and locations.

The result is an interactive Shiny dashboard featuring maps, plots, and data tables that let users explore Minnesota’s state parks in depth. Whether you’re planning a trip or researching park trends, this tool provides insight into the state’s natural spaces. 

**Data Sources:** [stateparks.com](http://stateparks.com), [Wikipedia](https://en.wikipedia.org/wiki/List_of_Minnesota_state_parks), [MN DNR GIS](https://gis.data.mn.gov/maps/a42128766db2447cb77a247d4a074173/about), [Joint Legacy Research Project](https://storymaps.arcgis.com/stories/8489886967dc48b58cb8f54e1a3c014f)

**Authors:** Mary Phillips, Kaisa Nilsson, Romina Amirian


Graphs
==========================

Column
-------------------------

This is a scatterplot of visitation numbers by year of park established, and the color of the points is by the season, either Fall, Spring, Summer, or Winter
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(y = total,
                      x = Year)) +
    geom_point() +
    labs(x = "Visitation (Car Counts)", 
         y = "Year Park Established",
         title = "Visitation by Year Park Established and Season")
})
```

We can also look at some other data. This scatterplot charts the number of acres of a park against the year it was established. Theres...not much of a relationship that we can see.
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(x = Acres,
                      y = Year)) +
    geom_point() +
    labs(x = "Acres", 
         y = "Year Park Established",
         title = "Acres by Park Established Year")
})
```


Maps - Individual Parks
==========================


```{r, label = "adding html labels", include = FALSE}
# joining polygons
parks_sf <- parks |>
  left_join(park_polygons, by = c("name" = "legislative_name"))

# adding html labels of name/description
parks_sf <- parks_sf |>
  mutate(labels1 = str_c("<p><strong>", name, "</strong></p>
<p><span style='font-weight: 400;'>", description, "</span></p>"))

# applying html
labels1 <- lapply(parks_sf$labels1, HTML)

# making the dataframe sf
parks_sf <- st_as_sf(parks_sf)

# checking class
class(parks_sf)
```

```{r}
renderLeaflet({
  leaflet(parks_sf) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 1,
    weight = 2,
    opacity = 1,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "red",
      bringToFront = TRUE),
    label = labels1,
    labelOptions = popupOptions( 
      style = list("font-weight" = "normal", padding = "2px 5px", color = "black", "width" = "450px", "white-space" = "normal"),
      textsize = "15px",
      direction = "auto")
  )
})
```

Maps - Counties
===================================

```{r, label = "lat long stuff", include = FALSE}
lat_long <- with_wiki |>
  mutate(long = str_replace_all(long, "^[^\\d]*", "-")) |>
  mutate(lat = as.numeric(lat), long = parse_number(long))
```

```{r, label = "setting up map", include = FALSE}
county_parks <- lat_long |>
  mutate(`County or Counties` = case_when(
    str_detect(`County or Counties`, "[a-z] and ") ~ str_replace_all(`County or Counties`, " and", ","),
    str_detect(`County or Counties`, "[a-z], and ") ~ str_replace_all(`County or Counties`, " and", ""),
    .default = `County or Counties`)) |>
  mutate(unpacked = str_split(`County or Counties`, ",")) |>
  filter(!is.na(unpacked)) |>
  unnest(cols = c(unpacked)) |>
  mutate(county = str_trim(unpacked)) |>
  #pivot_wider(names_from = county, values_from = n, values_fill = list(n = 0))
  group_by(county) |>
  mutate(num = n())
  
county_parks <- county_parks |>
  mutate(county = ifelse(county == "Murray County", "Murray", county))

county_polygons <- read_sf("Mn_County_Boundaries_Multipart.geojson")

county_maps <- county_parks |>
  right_join(county_polygons, by = c("county" = "countyname")) |>
  mutate(num = replace_na(num, 0))

pal <- colorFactor(c("white", "yellow2", "yellowgreen", "forestgreen", "cyan4", "blue4"), county_maps$num)

county_maps <- county_maps |>
  mutate(labels = str_c(county, " County: ", num, ifelse(num == 1, " state park", " state parks")),
         labels2 = name)

labels <- lapply(county_maps$labels, HTML)
labels2 <- lapply(county_maps$labels2, HTML)

county_maps <- st_as_sf(county_maps)
class(county_maps)
```

```{r}
leaflet(county_maps) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(num),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels2,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal, values = ~num, opacity = 0.7, title = NULL,
    position = "bottomright")
```