---
title: "state_parks_flexdashboard"
output: flexdashboard::flex_dashboard
runtime: shiny
---

```{r}
library(leaflet)
library(tidyverse)
library(sf)
library(readr)
```

```{r, label = "reading in csvs", include = FALSE}
numbers <-read_csv("numbers.csv")

with_wiki <- read_csv("wiki_parks_data.csv")

parks <- read_csv("parks_updated.csv")

park_polygons <- read_sf("Dnr_Stat_Plan_Areas_Prk.geojson") |>
  mutate(legislative_name = str_replace_all(legislative_name, "\\.", ""),
         legislative_name = ifelse(legislative_name == "St Croix State Park", "Saint Croix State Park", legislative_name))
```

About
==========================

Column {data-width = 200}
-------------------------

This is our website for MN stateparks.

Graphs
==========================

Column
-------------------------

This is a scatterplot of visitation numbers by year of park established, and the color of the points is by the season, either Fall, Spring, Summer, or Winter
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(x = total,
                      y = Year,
                      color = Season)) +
    geom_point() +
    labs(x = "Visitation (Car Counts)", 
         y = "Year Park Established",
         title = "Visitation by Year Park Established and Season")
})
```

We can also look at some other data. This scatterplot charts the number of acres of a park against the year it was established. Theres...not much of a relationship that we can see.
```{r echo = FALSE}
renderPlot({
  ggplot(numbers, aes(x = Acres,
                      y = Year)) +
    geom_point() +
    labs(x = "Acres", 
         y = "Year Park Established",
         title = "Acres by Park Established Year")
})
```


Maps - Individual Parks
==========================


```{r, label = "adding html labels", include = FALSE}
# joining polygons
parks_sf <- parks |>
  left_join(park_polygons, by = c("name" = "legislative_name"))

# adding html labels of name/description
parks_sf <- parks_sf |>
  mutate(labels = str_c("<p><strong>", name, "</strong></p>
<p><span style='font-weight: 400;'>", description, "</span></p>"))

# applying html
labels <- lapply(parks_sf$labels, HTML)

# making the dataframe sf
parks_sf <- st_as_sf(parks_sf)

# checking class
class(parks_sf)
```

```{r}
renderLeaflet({
  leaflet(parks_sf) |>
  setView(-93.8, 46.5, 6) |>
  addTiles() |>
  addPolygons(
    color = "darkred",
    fillColor = "darkred",
    fillOpacity = 1,
    weight = 2,
    opacity = 1,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "red",
      bringToFront = TRUE),
    label = labels,
    labelOptions = popupOptions( 
      style = list("font-weight" = "normal", padding = "2px 5px", color = "black", "width" = "450px", "white-space" = "normal"),
      textsize = "15px",
      direction = "auto")
  )
})
```

Maps - Counties
===================================

```{r, label = "lat long stuff", include = FALSE}
lat_long <- with_wiki |>
  mutate(long = str_replace_all(long, "^[^\\d]*", "-")) |>
  mutate(lat = as.numeric(lat), long = parse_number(long))
```

```{r, label = "setting up map", include = FALSE}
county_parks <- lat_long |>
  mutate(`County or Counties` = case_when(
    str_detect(`County or Counties`, "[a-z] and ") ~ str_replace_all(`County or Counties`, " and", ","),
    str_detect(`County or Counties`, "[a-z], and ") ~ str_replace_all(`County or Counties`, " and", ""),
    .default = `County or Counties`)) |>
  mutate(unpacked = str_split(`County or Counties`, ",")) |>
  filter(!is.na(unpacked)) |>
  unnest(cols = c(unpacked)) |>
  mutate(county = str_trim(unpacked)) |>
  #pivot_wider(names_from = county, values_from = n, values_fill = list(n = 0))
  group_by(county) |>
  mutate(num = n())
  
county_parks <- county_parks |>
  mutate(county = ifelse(county == "Murray County", "Murray", county))

county_polygons <- read_sf("Mn_County_Boundaries_Multipart.geojson")

county_maps <- county_parks |>
  right_join(county_polygons, by = c("county" = "countyname")) |>
  mutate(num = replace_na(num, 0))

pal <- colorFactor(c("white", "yellow2", "yellowgreen", "forestgreen", "cyan4", "blue4"), county_maps$num)

county_maps <- county_maps |>
  mutate(labels = str_c(county, " County: ", num, ifelse(num == 1, " state park", " state parks")),
         labels2 = name)

labels <- lapply(county_maps$labels, HTML)
labels2 <- lapply(county_maps$labels2, HTML)

county_maps <- st_as_sf(county_maps)
class(county_maps)
```

```{r}
leaflet(county_maps) |>
  setView(-93.8, 46.5, 7) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(num),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    ) |>
  addCircleMarkers(
    lng = ~long,
    lat = ~lat,
    radius = 2,
    color = "white",
    fillOpacity = 1,
    options = pathOptions(pane = "markerPane"),
    label = labels2,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px", color = "black"),
      textsize = "15px",
      direction = "auto")
  ) |>
  addLegend(pal = pal, values = ~num, opacity = 0.7, title = NULL,
    position = "bottomright")
```